---
title: "Tirgul 2 - EDA"
author: "Romi Goldner Kabeli"
date: "3/21/2023"
output:
  pdf_document: default
  html_document: default
---

We'll perform Exploratory Data Analysis ("EDA") on the Chocolates dataset.

First, load the libraries

```{r setup, include=FALSE}
library(tidyverse)
```
```{r setup, include=FALSE}
library(ggplot2)
library(ggrepel)
library(cowplot)
library(GGally)
library(ggplot2)
library(ggrepel)
library(cowplot)
```

then we load the data

```{r}

flavors <- read.csv("/Users/mashashopen/Downloads/flavors_of_cacao.csv", header = TRUE)

```



```{r}
# get dimensions, number of rows and columns
dim(flavors)
nrow(flavors) 
ncol(flavors)
```


JUST TO GET AN IDEA OF THE DATA. FOR EXAMPLE, THERE ARE ROWS WITHOUT "BEAN.TYPE" VALUE ETC

```{r}
# show first and last 6 (default) rows
head(flavors)
tail(flavors)

```
FULL LIST OF THE COLUMNS NAMES


```{r}
# column names
names(flavors)
```

STR - SHOWS US THE DATA AND THE TYPE OF EACH COLUMN
SUMMERY - USEFUL DATA ABOUT THE DATA VALUES IN EACH COLUMN

```{r}
# structure and summary of the data
str(flavors)
summary(flavors)
```


```{r}
# Convert Cocoa.Percent column to numeric type
flavors$Cocoa.Percent <- as.numeric(gsub("%","", flavors$Cocoa.Percent))
```




We can access specific values of a dataset using these commands:

```{r}

#names(flavors) <- make.names(names(flavors), unique = TRUE)



```


```{r}

# Distribution of Cocoa Percentages
ggplot(flavors, aes(x = Cocoa.Percent)) +
  geom_histogram(binwidth = 0.02, color = "black", fill = "blue") +
  labs(x = "Cocoa Percentage", y = "Count") +
  ggtitle("Distribution of Cocoa Percentages")

# Distribution of Ratings
ggplot(flavors, aes(x = Rating)) +
  geom_histogram(binwidth = 0.25, color = "black", fill = "orange") +
  labs(x = "Rating", y = "Count") +
  ggtitle("Distribution of Ratings")

# Top Chocolate Brands by Count
top_brands <- flavors %>% 
  count(Company...Maker.if.known.) %>% 
  top_n(10)

ggplot(top_brands, aes(x = reorder(Company...Maker.if.known., -n), y = n)) +
  geom_bar(stat = "identity", color = "black", fill = "green") +
  coord_flip() +
  labs(x = "Brand", y = "Count") +
  ggtitle("Top Chocolate Brands by Count")

# Top Chocolate Companies by Count
top_companies <- flavors %>% 
  count(Broad.Bean.Origin) %>% 
  top_n(10)

ggplot(top_companies, aes(x = reorder(Broad.Bean.Origin, -n), y = n)) +
  geom_bar(stat = "identity", color = "black", fill = "red") +
  coord_flip() +
  labs(x = "Bean Origin", y = "Count") +
  ggtitle("Top Beans' Origin by Count")

# Relationship between Cocoa Percentage and Rating
ggplot(flavors, aes(x = Cocoa.Percent, y = Rating)) +
  geom_point(color = "purple", alpha = 0.5) +
  labs(x = "Cocoa Percentage", y = "Rating") +
  ggtitle("Relationship between Cocoa Percentage and Rating") +
  geom_smooth(method = 'lm', se = FALSE, col = 'red')


# Top Chocolate Companies by Mean Rating
top_mean_ratings <- flavors %>% 
  group_by(Company...Maker.if.known.) %>% 
  summarize(mean_rating = mean(Rating)) %>% 
  top_n(10)

ggplot(top_mean_ratings, aes(x = reorder(Company...Maker.if.known., -mean_rating), y = mean_rating)) +
  geom_bar(stat = "identity", color = "black", fill = "brown") +
  coord_flip() +
  labs(x = "Company", y = "Mean Rating") +
  ggtitle("Top Chocolate Companies by Mean Rating")


```


## Data Variation

One important fact to know about our data is its variation, can you think of reasons why? 




```{r}
# creating boxplot
ggplot(data = flavors, mapping = aes(x = Rating, y = Cocoa.Percent)) + 
  geom_boxplot(mapping = aes(group = cut_width(Rating, 0.1)))

# Variance between two categorical variables
ggplot(data = flavors) +
  geom_count(mapping = aes(x = Company.Location, y = Bean.Type)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

Which plot is better? What do we know about these two variables in regard to their variation?

## Data Visualisation

```{r}
avg_ratings <- flavors %>% 
  group_by(Review.Date) %>% # Group by new date column
  summarise(avg_rating = mean(Rating), count_ratings = n()) # Summary stats

ggplot(avg_ratings, aes(x = Review.Date, y = avg_rating)) +
  geom_line() +
  labs(x = "Review Date", y = "Average Rating")

```




Now we can try to take a closer look at these values:

```{r}
unusual_values <- flavors %>% 
  filter(Rating < 2.75 | Rating > 4) %>% 
  select(Rating, Cocoa.Percent, Company.Location, Company...Maker.if.known., Bean.Type) %>%
  arrange(Cocoa.Percent)
unusual_values
```



## finding duplicated rows

```{r}

numeric_vars <- c(
  "Rating", "Review.Date", "Cocoa.Percent", "REF"
)
flavors_num <- flavors[, numeric_vars]

# Handle missing data
flavors_num[is.na(flavors_num)] <- 0


# Identify missing values
missing_values <- apply(flavors_num, 2, function(x) sum(is.na(x)))
flavors_num <- flavors_num[complete.cases(flavors_num), ]

# Remove duplicates
flavors_num <- unique(flavors_num)

# Check for outliers
boxplot(flavors_num)
flavors_num <- flavors_num[apply(flavors_num, 1, function(x) all(abs(scale(x)) < 3)), ]

```
CLASSIFICATION TO GOOD AND BAD CHOCLATES, BY RATING

```{r}

bad <- c(335, 439, 467, 556) # below 2
good <- c(88,80, 1292, 58) # above 4

flavors_num$classification <- "normal"
flavors_num$classification[match(good, rownames(flavors_num))] <- "good"
flavors_num$classification[match(bad, rownames(flavors_num))] <- "bad"

```


PCA

```{r}

# Perform PCA
pca_results <- prcomp(flavors_num[, -c(5)], center = TRUE, scale. = TRUE)

# Extract principal components
pc_data <- data.frame(
  PC1 = pca_results$x[, 1],
  PC2 = pca_results$x[, 2],
  classification = as.factor(flavors_num$classification)
)


# Visualize results

ggplot(pc_data, aes(x = PC1, y = PC2, col = classification, size = ifelse(classification == "normal", 1, 2))) +
geom_point() 
```



Using the PCA, we can now find chocolates that are similar to the good chocolates (based on rating).
In order to do that, we will look for the chocolates with similar values of the principal components as the good chocolates.

we can see the chocolates from Canada and Italy are approximately close to the good chocolates, and it may indicate that chocolate coming from these countries is good.



##################### IM NOT SURE WHERE ARE THE LABLES COMING FROM!!!!!!!#############################

```{r}
flavors_num_int <- which(pc_data$PC1 < 3 &
  pc_data$PC1 > 1.5 &
  pc_data$PC2 < 1 &
  pc_data$PC2 > -3)

pc_data$label2 <- ""
pc_data$label2[flavors_num_int] <- label[flavors_num_int]

ggplot(pc_data, aes(x = PC1, y = PC2, label = label2, col = classification, size = ifelse(classification == "normal", 1, 2))) +
  geom_point() +
  ggrepel::geom_text_repel(cex = 2.5, 
                           box.padding = 0.5, 
                           point.padding = 0.5, 
                           min.segment.length = 0.1,
                           force = 10,
                           max.iter = 1000)
```


```{r}
  library(Rtsne)


tsne_flavors_num <- Rtsne(flavors_num[,1:5],
  pca = FALSE, perplexity = 10,
  theta = 0.0
)

tsne_flavors_num <- data.frame(
   TSNE1 = tsne_flavors_num$Y[, 1],
   TSNE2 = tsne_flavors_num$Y[, 2],
   classification = flavors_num$classification
 )

```


```{r}
ggplot(tsne_flavors_num, aes(
 x = TSNE1, y = TSNE2,
 col = classification, size = ifelse(classification == "normal", 1, 2))) +
 geom_point()
```




```{r}
library(mlbench) 
library(caret)
library(randomForest)
library(ggcorrplot)
```


STRONG CORRELATION BETWEEN REF AND REVIEW.DATE (obvious...)
SOME CORRELATION BETWEEN RATING AND REF

```{r}

# calculate correlation matrix
correlationMatrix <- cor(flavors_num[-5])
# display the correlation matrix
ggcorrplot(correlationMatrix)
```









